---
title: "Detailed Usage Example"
output: rmarkdown::html_vignette
vignette: >
   %\VignetteIndexEntry{Vignette Title}
   %\VignetteEngine{knitr::rmarkdown}
   \usepackage[utf8]{inputenc}
---

```{r, include=FALSE}
devtools::load_all(".")
```

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
library(immunedeconv)
```

*Immunedeconv* ships with an example dataset containing purified immune cells extracted from several donors.
For demonstration purposes, let's create a subset of this dataset and estimate immune cell fractions.

```{r}
my_matrix = example_gene_expression_matrix[,c(1,2,12,13,19,20)]
```

Gene expression in rows:
```{r}
rownames(my_matrix)[1:5]
```

Samples in columns:
```{r}
colnames(my_matrix)
```

To estimate immune cell fractions, we simply have to invoke the `deconvolute` function. It requires the specification
of one of the following methods for deconvolution:

```{r}
deconvolution_methods
```


For this example, we use `quanTIseq`. As a result, we obtain a cell_type x sample data frame with cell-type scores for each sample.

```{r, message=FALSE, warning=FALSE}
res_quantiseq = deconvolute(my_matrix, "quantiseq")
```


QuanTIseq generates scores that can be interpreted as a cell-type fraction. Let's visualize
the results as a stacked bar chart with tidyverse/ggplot2.
```{r, fig.height=3, fig.width=8}
res_quantiseq %>%
  gather(sample, fraction, -cell_type) %>%
  ggplot(aes(x=sample, y=fraction, fill=cell_type)) +
    geom_bar(stat='identity') +
    coord_flip() +
    scale_fill_brewer(palette="Paired")
```

We observe that the correct cell-types are correctly recovered, however we observe small fractions of NK cells and CD4+ predicted to be present in CD8+ T cells and considerable fractions of CD8+ and regulatory T cells in NK cells. This can either be due to the fact that the purification of the samples has not worked reliably or due to spillover-effects between closely-related cell-types (see our benchmark).


Let's now apply MCP-counter to the same dataset.
```{r, message=FALSE, warning=FALSE}
res_mcp_counter = deconvolute(my_matrix, "mcp_counter")
```
MCP-counter provides scores in arbitrary units that are only comparable between samples, but not between
cell-types. The visualisation as bar-chart suggests the scores to be cell-type fractions and is, therefore, unsuitable.
Instead, we use ggplot to visualize the scores per-cell type, allowing for a relative comparison between samples.

```{r, fig.height=5, fig.width=10}
res_mcp_counter %>%
  gather(sample, score, -cell_type) %>%
  ggplot(aes(x=sample, y=score, color=cell_type)) +
    geom_point() +
    facet_wrap(~cell_type, scales="free_y", nrow=2) +
    scale_color_brewer(palette="Paired", guide=FALSE) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

With the scores being in arbitrary units, the results are not useful for judging if a cell type is present in the sample, or not. However, the score correctly reflects the presence of NK cells and CD8+ T cells in the corresponding samples.

