---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
  word_document: default
---

```{r setup, include=FALSE}
library(dplyr)
library(immunedeconv)
library(ggplot2)
library(tidyr)
```

*Immunedeconv* ships with an example dataset containing purified immune cells extracted from several donors. 
For demonstration purposes, let's create a subset of this dataset and estimate immune cell fractions. 

```{r}
my_matrix = example_gene_expression_matrix[,c(1,2,12,13,19,20)]

# Gene expression in rows
rownames(my_matrix)[1:5]

# samples in columns
colnames(my_matrix)
```

To estimate immune cell fractions, we simply have to invoke the `deconvolute` function. It requires the specification 
of one of the following methods for deconvolution:

```{r}
deconvolution_methods
```

For this example, we use `quanTIseq`. As a result, we obtain a cell_type x sample data frame with cell-type scores for each sample. 

```{r}
res_quantiseq = deconvolute(my_matrix, "quantiseq")
```


QuanTIseq generates scores that can be interpreted as a cell-type fraction. Let's visualize 
the results as a stacked bar chart with tidyverse/ggplot2. 
```{r, fig.height=3, fig.width=8}
res_quantiseq %>% 
  gather(sample, fraction, -cell_type) %>% 
  ggplot(aes(x=sample, y=fraction, fill=cell_type)) + 
    geom_bar(stat='identity') +
    coord_flip() +
    scale_fill_brewer(palette="Paired")

ggsave("/tmp/fig1.pdf")
```

We observe that the correct cell-types are correctly recovered, however we observe small fractions of NK cells and CD4+ predicted to be present in CD8+ T cells and considerable fractions of CD8+ and regulatory T cells in NK cells. This can either be due to the fact that the purification of the samples has not worked reliably or due to spillover-effects between closely-related cell-types (see our benchmark).  


Let's now apply MCP-counter to the same dataset. 
```{r}
res_mcp_counter = deconvolute(my_matrix, "mcp_counter")
```
MCP-counter provides scores in arbitrary units that are only comparable between samples, but not between 
cell-types. The visualisation as bar-chart suggests the scores to be cell-type fractions and is, therefore, unsuitable.
Instead, we use ggplot to visualize the scores per-cell type, allowing for a relative comparison between samples. 

```{r, fig.height=5, fig.width=10}
res_mcp_counter %>% 
  gather(sample, score, -cell_type) %>% 
  ggplot(aes(x=sample, y=score, color=cell_type)) + 
    geom_point() + 
    facet_wrap(~cell_type, scales="free_y", nrow=2) + 
    scale_color_brewer(palette="Paired", guide=FALSE) + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggsave("/tmp/fig2.pdf")
```

With the scores being in arbitrary units, the results are not useful for judging if a cell type is present in the sample, or not. However, the score correctly reflects the presence of NK cells and CD8+ T cells in the corresponding samples. 

