---
title: "Detailed Usage Example"
output: rmarkdown::html_vignette
vignette: >
   %\VignetteIndexEntry{Detailed Exampe}
   %\VignetteEngine{knitr::rmarkdown}
   \usepackage[utf8]{inputenc}
---

```{r, include=FALSE}
devtools::load_all(".")
```

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
library(immunedeconv)
```

*Immunedeconv* ships with an example dataset with samples from four patients with metastatic melanoma published in @EPIC2017. 
It is available from `immunedeconv::dataset_racle`. It contains a gene expression matrix (`dataset_racle$expr_mat`) generated using bulk RNA-seq and 'gold standard' estimates of immune cell contents profiled with FACS (`dataset_racle$ref`). We are going to use the bulk RNA-seq data to run the deconvolution methods and will compare the results to the FACS data later on.   

The gene expression data is a matrix with HGNC symbols in rows and samples in columns:
The dataset is a matrix 
```{r}
# show the first 5 lines of the gene expression matrix
knitr::kable(dataset_racle$expr_mat[1:5, ])
```



To estimate immune cell fractions, we simply have to invoke the `deconvolute` function. It requires the specification
of one of the following methods for deconvolution:

```{r}
deconvolution_methods
```


For this example, we use `quanTIseq`. As a result, we obtain a cell_type x sample data frame with cell-type scores for each sample.

```{r, message=FALSE, warning=FALSE}
res_quantiseq = deconvolute(dataset_racle$expr_mat, "quantiseq", tumor = TRUE)
```


QuanTIseq generates scores that can be interpreted as a cell-type fraction. Let's visualize
the results as a stacked bar chart with tidyverse/ggplot2.
```{r, fig.height=3, fig.width=8}
res_quantiseq %>%
  gather(sample, fraction, -cell_type) %>%
  # plot as stacked bar chart
  ggplot(aes(x=sample, y=fraction, fill=cell_type)) +
    geom_bar(stat='identity') +
    coord_flip() +
    scale_fill_brewer(palette="Paired") + 
    scale_x_discrete(limits = rev(levels(res_quantiseq)))
```

We observe that

* the scores of the first two samples sum to 1, while the last two don't. #TODO other cells
* the first two samples contain a large quantity of B cells and CD4+ T cells. 



Let's now apply MCP-counter to the same dataset.
```{r, message=FALSE, warning=FALSE}
res_mcp_counter = deconvolute(dataset_racle$expr_mat, "mcp_counter")
```
MCP-counter provides scores in arbitrary units that are only comparable between samples, but not between
cell-types. The visualisation as bar-chart suggests the scores to be cell-type fractions and is, therefore, unsuitable.
Instead, we use ggplot to visualize the scores per-cell type, allowing for a relative comparison between samples.

```{r, fig.height=7, fig.width=7}
res_mcp_counter %>%
  gather(sample, score, -cell_type) %>%
  ggplot(aes(x=sample, y=score, color=cell_type)) +
    geom_point(size=4) +
    facet_wrap(~cell_type, scales="free_x", ncol=3) +
    scale_color_brewer(palette="Paired", guide=FALSE) +
    coord_flip() + 
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

With the scores being in arbitrary units, the results are not useful for judging if a cell type is present in the sample, or not. However, the score correctly reflects the presence of NK cells and CD8+ T cells in the corresponding samples.


## TODO add comparison with FACS data. 
